Effects of CHILDREN's programs

### 2020-02-27 paper.Rnw
xxx

@@

% working with knitR and LaTeX
% https://kbroman.org/knitr_knutshell/pages/latex.html
% Command line
% R -e 'library(knitr);knit("knitr_example.Rnw")'

@@

\documentclass[12pt, a4paper, titlepage]{article}
\usepackage{lscape}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage[backend=biber, bibstyle=apa, citestyle=authoryear]{biblatex}
\usepackage{floatrow}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}
\usepackage{rotating}
% \usepackage[authoryear]{natbib}
% \bibliographystyle{apalike}
\usepackage{mathtools}
\usepackage[onehalfspacing]{setspace}
\usepackage[left = 2.5cm, top = 2.5cm, right = 2.5cm, bottom = 2.0cm]{geometry}

@@

@@

\begin{figure}
  \caption{Dispersion in productivity levels}
  \label{LineDisp}

<<echo=FALSE>>=
figureLineDisp = readRDS("./ANALYSIS/GRAPHS/LineDisp.Rds")
print(figureLineDisp)
@

\floatfoot{Note: Time series plots of approximate deviations of the three annual dispersion measures from their trends in percent. A shows the full sample, B the non-durable manufacturing sector, and C the durable manufacturing sector. After taking the natural logarithm of the dispersion measures defined in equation 1, I have isolated their cyclical components with an HP-100 filter. The shaded bars represent recessions as defined by the NBER. The year ticks refer to January 1. The dispersion measures take as their date the middle of the year, July 2. Compare Kehrig (2015), Figure 1.}
\end{figure}

Figure \ref{LineDisp}

@@

<<mealsTripsSubsidyRegressions, echo=FALSE, results='asis'>>=
# , message=FALSE, warning=FALSE
# first, read in all Rds objects that you need
# 
mealsNo_sub.lm = readRDS("./ANALYSIS/Tables/mealsNo_sub.lm.Rds")
tripsNo_sub.lm = readRDS("./ANALYSIS/Tables/tripsNo_sub.lm.Rds")
#  
texreg(list(mealsNo_sub.lm, tripsNo_sub.lm))
@

@@

@@

<<echo=FALSE>>=
figureDaytodayskillsSelfworth = readRDS("./ANALYSIS/GRAPHS/PAPER/plot_dayToDaySkills_selfworth.Rds")
print(figureDaytodayskillsSelfworth)
@

@@

@@

fundamental_dynamics <- xtable(fundamental_dynamics,
                               digits = c(0, 0, 1, 1, 0, 0),
          caption = "Summary Statistics",
          align = "lmmmmm", 
          label = "fundamentalDynamics"
          )
print(fundamental_dynamics, include.rownames = FALSE)
@
% look up official names of price indices in English

@@

Figure \ref{Health_Plots} gives an overview of the relationship between the healthy food criterion and each of the three health-related variables. The x-axis plots the index for a healthy diet. The y-axis displays the share of organizarions in each category of the health outcome. The possible values are: all(coded as 4), most (3), some (2), few (1), and none (coded as 0). For example, if an organization says that most beneficiaries are healthier, then this would be coded as 3.

We use standardized versions of these variables as outcomes in simple linear models with a standardized version of the healthy food criterion as predictor in each.
If we estimated these models with ordinary least squares, we would ascribe the same weight to an organization where a thousand beneficiaries regularly eat as to one where only ten people do so. To control for this difference in size, we fit the models with weighted least squares, using the number of beneficiaries as weights.

@@

In the following equation we use standardized variables in a standard OLS approach. The results are as well compared to the output with the imputed dataset, where we also use scaled coefficents.

@@

<<echo=FALSE, results='asis', message=FALSE>>=

@@

\subsection{A1: Cumulative Odds Regression}

As an option to run an regression on the original ordinal variables, we use the vglm function from the VGAM package...

Use from zwichenpräsentation

\begin{equation}
\label{ModelProdu}
\ln y_{it} = \frac{}{}  \sum
\end{equation}

<<echo=FALSE, results='asis', message=FALSE>>=

@@

stargazer(lessIllOdds@coef3, title= "Propodss Regression Results: Less Ill", align = TRUE)

@
<<echo=FALSE, results='asis', message=FALSE>>=

@@

stargazer(dietaryOdds@coef3, title= "Propodss Regression Results: Dietary Knowledge", align = TRUE)
@
<<echo=FALSE, results='asis', message=FALSE>>=

@@

stargazer(appreciateOdds@coef3, title= "Propodss Regression Results: Appreciate Healthy", align = TRUE)
@

<<echo=FALSE, results='asis', message=FALSE>>=

@@

stargazer(selfworthOdds@coef3, title= "Propodss Regression Results: Equality of opportunities, selfworth", align = TRUE)
@

<<echo=FALSE, results='asis', message=FALSE>>=

@@

stargazer(dayToDayOdds@coef3, title= "Propodss Regression Results: Equality of opportunities, everyday expertise", align = TRUE)
@

\subsection{A2: Partition}

In addition to the factor analysis described in section 6, we would like to introduce to a another dimensionality reduction method called partition.

QUOTES FROM PAPER, describe method

In the following a threshold of 0.4 is used, meaning that the reduced variable consists of variables which explain each other to at least 40 percent. It might be meaningful to decide to use only one of the variables a reduced variable consists of or a summarizing one for future surveys to avoid redundancy. Table x shows the obtained results of the dimensionality reduction and displays x variables which haven't been reduced, and x reduced ones.

@@

xtable(partition_meals,
                               #digits = c(0, 0, 1, 1, 0, 0),
          caption = "Partition of outcomes, Meals"
          #align = "lmmm"
          )
#print(partition_meals, include.rownames = FALSE)

xtable(partition_trips,
                               #digits = c(0, 0, 1, 1, 0, 0),
          caption = "Partition of outcomes, Trips"
          #align = "lmmm"
          )
print(partition_meals, partition_trips, include.rownames = FALSE)
@
\subsection{A3: OLS Regressions}

#normal weights selfworth& subsidiy
<<eqaulity_extra_reg, echo=FALSE, results='asis', message=FALSE, warning=FALSE>>=

>>>>

% instead of Compile PDF, go into Terminal and enter 
% Rscript -e "library(knitr); knit('paper.Rnw')"
% latexmk -pdf paper

@@

% working with knitR and LaTeX
% https://kbroman.org/knitr_knutshell/pages/latex.html

@@

\documentclass[12pt, a4paper, titlepage]{article}
\usepackage{lscape}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage[backend=biber, bibstyle=apa, citestyle=authoryear]{biblatex}
\usepackage{floatrow}
\usepackage{booktabs}
% \usepackage{bookmark}
\usepackage{rotating}
% \usepackage[authoryear]{natbib}
% \bibliographystyle{apalike}
\usepackage{mathtools}
% \usepackage[onehalfspacing]{setspace}
\usepackage[left = 2.5cm, top = 2.5cm, right = 2.5cm, bottom = 2.0cm]{geometry}
\usepackage[hidelinks]{hyperref}

@@

Citing\\

If you want to cite \textcite{Millenstein.2020} inline, use this.
If you want to cite \parencite[p.4]{Millenstein.2020} inline with a page number, use this.
If you want to cite \citeauthor{Millenstein.2020} only as authors, use this.

@@

% Figure\\
% 
% \begin{figure}
%   \caption{Dispersion in productivity levels}
%   \label{LineDisp}
% 
% <<echo=FALSE>>=
% figureLineDisp = readRDS("./ANALYSIS/GRAPHS/LineDisp.Rds")
% print(figureLineDisp)
% @
% 
% \floatfoot{Note: Time series plots of approximate deviations of the three annual dispersion measures from their trends in percent. A shows the full sample, B the non-durable manufacturing sector, and C the durable manufacturing sector. After taking the natural logarithm of the dispersion measures defined in equation 1, I have isolated their cyclical components with an HP-100 filter. The shaded bars represent recessions as defined by the NBER. The year ticks refer to January 1. The dispersion measures take as their date the middle of the year, July 2. Compare Kehrig (2015), Figure 1.}
% \end{figure}

% Figure \ref{LineDisp}

@@

<<mealsTripsSubsidyRegressions, echo=FALSE, results='asis'>>=
# , message=FALSE, warning=FALSE
# first, read in all Rds objects that you need
# 
mealsNo_sub.lm = readRDS("./ANALYSIS/Tables/mealsNo_sub.lm.Rds")
tripsNo_sub.lm = readRDS("./ANALYSIS/Tables/tripsNo_sub.lm.Rds")
#  
texreg(list(mealsNo_sub.lm, tripsNo_sub.lm), label = "mealsTripsSubsidyRegressions")
@

@@

\subsection{Previous research}

The previous research shows, that the childhood and youth is central in development of character traits and future opportunities. Factors like education, stable family conditions, social contacts and mobility shape children and adolescents for their further life. Heckmann and Carneiro (2003) show that especially early family factors determine the most of gaps in highschool attendance. Moreover, Deckers et. al. (2015) found by experiments in Germany, that children of higher educated parents are signficantly more patient. Children from parents with lower income and lower educational attainment typically have a lower weight at birth and are born in a later week of gestation, which are indicators for good initial conditions (Deckers et. al. 2015, Case et. al. 2002). Children that grow up in families with a low socio-economic status are often less mobile and more likely to keep to themselves. This could have a negative influence on their character development. For example, Chetty et. al (2012) investigate for the US, that areas with more mobility are highly correlated with better primary schools, greater social capital and more stable family conditions. Furthermore, the result of Deckers et. al. (2020) provides causal evidence on the effect of social environment on prosocial attitudes. Being supported by a mentor could have a significant and persistant increase of elementary school children´s prosociality, in response to prosocial role models and intense social interactions. As investigated by Heckmann et. al. (2010) for a preschool education program in the US, investment in the improvement of the childhood conditions could have a high rate of return, even by controlling for possible distortions.

@@

<<DaytodayskillsSelfworth, echo=FALSE>>=
figureDaytodayskillsSelfworth = readRDS("./ANALYSIS/GRAPHS/PAPER/plot_dayToDaySkills_selfworth.Rds")
print(figureDaytodayskillsSelfworth)
@

@@

# <<FundamentalDynamics, echo=FALSE, results='asis', message=FALSE, warning=FALSE>>=
# 
# # first, read in all Rds objects that you need
# 
# fundamental_dynamics = readRDS("./ANALYSIS/Tables/fundamental_dynamics.Rds")
#  
# colnames(fundamental_dynamics) <- c("Year", "Beneficiaries, Meals", "Beneficiaries, Trips", "Organizations, Meals", "Organizations, Trips")
# 
# fundamental_dynamics <- xtable(fundamental_dynamics,
#                                digits = c(0, 0, 1, 1, 0, 0),
#           caption = "Summary Statistics",
#           align = "lmmmmm", 
#           label = "fundamentalDynamics"
#           )
# print(fundamental_dynamics, include.rownames = FALSE)
# @

@@

fundamental_dynamics <- stargazer(fundamental_dynamics,
                                  summary = FALSE,
                               # digits = c(0, 0, 1, 1, 0, 0),
          caption = "Summary Statistics",
          # align = "lmmmmm", 
          label = "fundamentalDynamics"
          )
print(fundamental_dynamics, include.rownames = FALSE)
@
% look up official names of price indices in English

@@

Figure \ref{Health_Plots} gives an overview of the relationship between the healthy food criterion and each of the three health-related variables. The x-axis displays the index for a healthy diet. The y-axis displays the share of organizarions in each category of the health outcome. The possible values are: all, most, some, few and none. For example, if an organization says that most beneficiaries are healthier, then this would be coded as most. The highest amount of healthy criteria to be fulfilled is x.

All plots of Figure \ref{Health_Plots} show that most organizations suffice about 5 to 10 healthy food criteria. The largest part of criterion values are associated with most beneficiaries being less Ill, having a growing appreciation for a healthy diet or having increased their knowledge about what constituted a healthy diet. It secondly becomes visible that the healthy criterion only exchanges with a growing appreciation for a healthy diet or an increasing dietary knowlege for all beneficiaries (which is the second most stated criterion in the two bottom plots), but not with all beneficaries being more healthy. However, for the variable less ill, the least data is availabe. Overall, the criteria, even a few, seem to have a positive relationship with beneficiaries health relevant variables. This connection is empiricaly examined in regressions \ref{HealthRegressions-LessIll}, \ref{HealthRegressions-DietaryKnowledge} and \ref{HealthRegressions-AppreciateHealthy}.

@@

We use standardized versions of the health related variables as outcomes in simple linear models with a standardized version of the healthy food criterion as predictor in each.
By estimating these models with ordinary least squares, we ascribe the same weight to an organization where a thousand beneficiaries regularly eat as to one where only ten people do so. To control for this difference in size, we additionally fit the models with weighted least squares, using the number of beneficiaries as weights.

In Table \ref{HealthRegressions-LessIll}, after weighting the regression for the original data set, the coefficent rises slightly but is becoming less statistically significant. An increase of the healthy food criterion by one standard deviation is associated with an increase in beneficiaris being more healthy by 0.35 standard deviations. Regressing with the imputed dataset, results decrease by about 0.1. Using weighted least squares additionaly results in a slighty decreased coefficent, which is not significant anymore.

@@

<<selfworth_tripsProgram, echo=FALSE, results='asis', message=FALSE>>=

@@

\subsection{A1: Expanded Health Regression}

In addition to table \ref{HealthRegressions-LessIll} table \ref{expandLessIll} presents the results of an OLS approach with scaled outcomes as well as a WLS approach with scaled outcomes with following control variables: to which extent an organization uses regional products, since when an organization is part of the Lunch program, the real subsidy per beneficiary and the state an organization is located.

% #normal weights selfworth& subsidiy
% <<eqaulity_extra_reg, echo=FALSE, results='asis', message=FALSE, warning=FALSE>>=
% 
% lm_s = readRDS("./ANALYSIS/Tables/selfworth.lm.Rds") #weighted
% lm_d = readRDS("./ANALYSIS/Tables/dayToDaySkills.lm.Rds") #weighted
% ts = readRDS("./ANALYSIS/Tables/selfworthTrips.lm.Rds")
% td = readRDS("./ANALYSIS/Tables/dayToDaySkillsTrips.lm.Rds")
% sI = readRDS("./ANALYSIS/Tables/selfworth_IM.lm.Rds") #weighted
% dI = readRDS("./ANALYSIS/Tables/dayToDaySkills_IM.lm.Rds") #weighted
% tsI = readRDS("./ANALYSIS/Tables/selfworthTrips_IM.lm.Rds")
% tdI = readRDS("./ANALYSIS/Tables/dayToDaySkillsTrips_IM.lm.Rds")
% @

<<lessIll_extra_reg, echo=FALSE, results='asis', message=FALSE, warning=FALSE>>=
#expanded model less ill

lm_elDGE = readRDS("./ANALYSIS/Tables/Expand_LessIll_scaled.Rds")
lm_elDGEw = readRDS("./ANALYSIS/Tables/Expand_LessIll_scaled_weighted.Rds")
elDGE_i = readRDS("./ANALYSIS/Tables/Expand_LessIll_IM_scaled.lm.Rds")
elDGE_iw = readRDS("./ANALYSIS/Tables/Expand_LessIll_IM_scaled_weighted.lm.Rds")

texreg(list(lm_elDGE, lm_elDGEw, elDGE_i, elDGE_iw), caption = "Regression Results: Less ill expanded model", custom.model.names = c("OLS", "WLS", "OLS Impute","WLS Impute"), scalebox = 0.8, label = 'expandLessIll')
@

\subsection{A2: Cumulative Odds Regression}

As an option to run a regression on the original ordinal variables, we use the proportional odds version of the cumulative logit from the VGAM package. Equation \ref{PropoddsModel} describes the model setup:

\begin{equation}
\label{PropoddsModel}
  P(Y_i \leq \ r) = \frac{exp(y_0_r + x_i^Ty)}{1+exp(y_0_r + x_i^Ty)}    
\end{equation}

To demonstrate the results of this method, we show in the tables \ref{lessIllOdds}, \ref{dietaryOdds}, \ref{appreciateOdds} the regressions of the health related variables on the healthy meals criterion and in the tables \ref{selfworthOdds} and \ref{dayToDayOdds} the regressions of the real meals subsidy on selfworth and everyday expertise.

As an example of interpreting such models, we consider the model of table \ref{lessIllOdds}: An increase of the healthy meals criterion by 1 is associated with an increase in chances, that a proportion of a maximum of r beneficiaries is healthier in relation to that a proportion of more than r beneficiaries are healthier, by the factor of exp(-0.29127) = 0.75.

A summary of the remaining models:

\begin{itemize}
  \item{Dietary Knowledge: exp(-0.089) = 0.91}
  \item{Appreciate Healthy: exp(-0199) = 0.82}
  \item{Selfworth: exp(-0.00001) = 1}
  \item{Everyday Expertise: (-0.00003) = 1} 
\end{itemize}

<<lessIllOdds, echo=FALSE, results='asis', message=FALSE>>=

@@

stargazer(lessIllOdds@coef3, title= "Propodss Regression Results: Less Ill", align = TRUE, label = "lessIllOdds")

<<dietaryOdds, echo=FALSE, results='asis', message=FALSE>>=

@@

stargazer(dietaryOdds@coef3, title= "Propodss Regression Results: Dietary Knowledge", align = TRUE, label = "dietaryOdds")
@

<<appreciateOdds, echo=FALSE, results='asis', message=FALSE>>=

@@

stargazer(appreciateOdds@coef3, title= "Propodss Regression Results: Appreciate Healthy", align = TRUE, label = "appreciateOdds")
@

<<selfworthOdds, echo=FALSE, results='asis', message=FALSE>>=

@@

stargazer(selfworthOdds@coef3, title= "Propodss Regression Results: Selfworth", align = TRUE, label = "selfworthOdds")
@

<<dayToDayOdds, echo=FALSE, results='asis', message=FALSE>>=

@@

stargazer(dayToDayOdds@coef3, title= "Propodss Regression Results: Everyday expertise", align = TRUE, label = "dayToDayOdds")
@

\subsection{A3: Partition}

In addition to the factor analysis which is described in section 6, we would like to introduce to a another dimensionality reduction method called partition. This method was developped by Millenstein et. al in the context of genome analysis.

According to the researchers, potential benefits of reducing dimensionality are reduced computational demands, reduced multiple-testing burden, reduced noise, as well as better-behaved data \cite{} (p.677).

The partition approach partitions data into subsets of related features. It then  summarizes each subset into one new feature. By that, it defines a surjective mapping \cite{} (p.676). A function g is implemented using the arithmetic mean. In contrast to methods like the factor analysis, if a feature is identified as related to a response of interest, the exact subset of features that are implicated to the surjective mapping is known without ambiguity. What is as well different is the information loss constraint. It lets the extent of dependences among features on a local level guide the extent to which the number of features is reduced.\cite{}(p.677)

In the following a threshold of 0.4 is used, meaning that the reduced variable consists of variables which explain each other to at least 40 percent. It might be meaningful to decide to use only one of the featured variables in a reduced variable (or a summarizing one) for future surveys to avoid redundancy.

Table \ref{partitionmeals} shows the obtained results of the dimensionality reduction for the variables of the Lunch program. It displays 10 variables which haven't been reduced, and 5 reduced variables with a number of featured variables.

Table \ref{partitiontrips} shows the partitioned results for the variables of the Trips program. Here, 12 variables haven't been reduced, and a number of featured variables are summarized to 3 reduced variables.

@@

stargazer(partition_meals,
                               #digits = c(0, 0, 1, 1, 0, 0),
          caption = "Partition of outcomes, Meals", 
          label = 'partitionmeals'
          #align = "lmmm"
          )
#print(partition_meals, include.rownames = FALSE)

stargazer(partition_trips,
                               #digits = c(0, 0, 1, 1, 0, 0),
          caption = "Partition of outcomes, Trips",
          label = 'partitiontrips'
          #align = "lmmm"
          )
print(partition_meals, partition_trips, include.rownames = FALSE)
@

\subsection{A4: OLS Regressions}

#normal weights selfworth& subsidiy
<<equality_extra_reg, echo=FALSE, results='asis', message=FALSE, warning=FALSE>>=
